---
title: "project_data"
author: "Josh Irwin"
date: "`r Sys.Date()`"
output: 
  html_document:
    TOC: TRUE
    number_sections: TRUE
    code_folding: show
    code_download: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Election Data

## 2000 Votes
```{r}
html_contents_votes2000 <- rvest::read_html("https://en.wikipedia.org/wiki/2000_United_States_presidential_election")
contents_votes2000 <- rvest::html_elements(x = html_contents_votes2000, css = "#mw-content-text > div.mw-content-ltr.mw-parser-output > div:nth-child(134) > table")
table_votes2000 <- rvest::html_table(html_contents_votes2000, header = 1)
magrittr::extract2(table_votes2000, 24) -> votes_2000

#remove unnecessary columns and rows
votes_2000 <- votes_2000[-c(1, 10, 22, 23, 32, 33, 34, 58), c(1, 3, 4, 6, 7, 9 ,10)]

#change column names
colnames(votes_2000)[1] = "State"
colnames(votes_2000)[2] = "Bush #"
colnames(votes_2000)[3] = "Bush %"
colnames(votes_2000)[4] = "Democratic #"
colnames(votes_2000)[5] = "Democratic %"
colnames(votes_2000)[6] = "Nader #"
colnames(votes_2000)[7] = "Nader %"

#reformat numbers
votes_2000$`Bush #` <- gsub(",", "", votes_2000$`Bush #`)
votes_2000$`Bush %` <- gsub("%", "", votes_2000$`Bush %`)
votes_2000$`Democratic #` <- gsub(",", "", votes_2000$`Democratic #`)
votes_2000$`Democratic %` <- gsub("%", "", votes_2000$`Democratic %`)
votes_2000$`Nader #` <- gsub(",", "", votes_2000$`Nader #`)
votes_2000$`Nader %` <- gsub("%", "", votes_2000$`Nader %`)

#change column data type
votes_2000$`Bush #` <- as.numeric(votes_2000$`Bush #`)
votes_2000$`Bush %` <- as.numeric(votes_2000$`Bush %`)
votes_2000$`Democratic #` <- as.numeric(votes_2000$`Democratic #`)
votes_2000$`Democratic %` <- as.numeric(votes_2000$`Democratic %`)
votes_2000$`Nader #` <- as.numeric(votes_2000$`Nader #`)
votes_2000$`Nader %` <- as.numeric(votes_2000$`Nader %`)

#percent reformatting
votes_2000$`Bush %` = votes_2000$`Bush %` / 100
votes_2000$`Democratic %` = votes_2000$`Democratic %` / 100
votes_2000$`Nader %` = votes_2000$`Nader %` / 100

#add election year column
votes_2000$ElectionYear <- 2000
```

## 2004 Votes
```{r}
html_contents_votes2004 <- rvest::read_html("https://en.wikipedia.org/wiki/2004_United_States_presidential_election")
contents_votes2004 <- rvest::html_elements(x = html_contents_votes2004, css = "#mw-content-text > div.mw-content-ltr.mw-parser-output > table.wikitable.sortable.jquery-tablesorter")
table_votes2004 <- rvest::html_table(html_contents_votes2004, header = 1)
magrittr::extract2(table_votes2004, 19) -> votes_2004

#remove unnecessary columns and rows
votes_2004 <- votes_2004[-c(1, 10, 22, 23, 32, 33, 34, 58), c(1, 3, 4, 6, 7, 9 ,10)]

#change column names
colnames(votes_2004)[1] = "State"
colnames(votes_2004)[2] = "Bush #"
colnames(votes_2004)[3] = "Bush %"
colnames(votes_2004)[4] = "Democratic #"
colnames(votes_2004)[5] = "Democratic %"
colnames(votes_2004)[6] = "Nader #"
colnames(votes_2004)[7] = "Nader %"

#reformat numbers
votes_2004$`Bush #` <- gsub(",", "", votes_2004$`Bush #`)
votes_2004$`Bush %` <- gsub("%", "", votes_2004$`Bush %`)
votes_2004$`Democratic #` <- gsub(",", "", votes_2004$`Democratic #`)
votes_2004$`Democratic %` <- gsub("%", "", votes_2004$`Democratic %`)
votes_2004$`Nader #` <- gsub(",", "", votes_2004$`Nader #`)
votes_2004$`Nader %` <- gsub("%", "", votes_2004$`Nader %`)

#change column data type
votes_2004$`Bush #` <- as.numeric(votes_2004$`Bush #`)
votes_2004$`Bush %` <- as.numeric(votes_2004$`Bush %`)
votes_2004$`Democratic #` <- as.numeric(votes_2004$`Democratic #`)
votes_2004$`Democratic %` <- as.numeric(votes_2004$`Democratic %`)
votes_2004$`Nader #` <- as.numeric(votes_2004$`Nader #`)
votes_2004$`Nader %` <- as.numeric(votes_2004$`Nader %`)

#percent reformatting
votes_2004$`Bush %` = votes_2004$`Bush %` / 100
votes_2004$`Democratic %` = votes_2004$`Democratic %` / 100
votes_2004$`Nader %` = votes_2004$`Nader %` / 100

#add election year column
votes_2004$ElectionYear <-  2004
```

## Combine Votes Data Frames
```{r}
votes <- rbind(votes_2000, votes_2004)
dplyr::glimpse(votes)
```

## Index Column
```{r}
votes$Index <- 0
for (i in 1:100) {
  votes$Index[i] <- i
  i + 1
}
```


## Write to Excel
```{r}
projectdata <- openxlsx::loadWorkbook("projectdata.xlsx")
openxlsx::addWorksheet(projectdata, "Votes")
openxlsx::writeData(projectdata, "Votes", votes)
openxlsx::saveWorkbook(projectdata, "projectdata.xlsx", overwrite = TRUE)
```

# Economic Data

## GDP Data
```{r}
gdp_data <- tidyquant::tq_get("GDP", get = "economic.data", from = "2000-01-01", to = "2009-01-30")
real_gdp_data <- tidyquant::tq_get("GDPC1", get = "economic.data", from = "2000-01-01", to = "2009-01-30")
```

## Combine Data Frames
```{r}
gdp <- rbind(gdp_data, real_gdp_data)
```

## Index Column
```{r}
nrow(gdp)
gdp$Index <- 0
for (i in 493:567) {
  gdp$Index[i - 493] <- i
  i + 1
}
```

## Relabel Symbols
```{r}
gdp$symbol[gdp$symbol == "GDP"] <- "Nominal"
gdp$symbol[gdp$symbol == "GDPC1"] <- "Real"
```


## Write to Excel
```{r}
openxlsx::addWorksheet(projectdata, "GDP")
openxlsx::writeData(projectdata, "GDP", gdp)
openxlsx::saveWorkbook(projectdata, "projectdata.xlsx", overwrite = TRUE)
```

# Armed Forces Data

## Armed Forces Data
```{r}
library(WDI)
library(tidyverse)

datTOT1 = WDI::WDI(indicator='MS.MIL.TOTL.P1', country=c('RU','CN','US','BRA','ESP','DEU','IND','ZAF'), start=1998, end=2011)
datTOT2 = WDI::WDI(indicator='SP.POP.TOTL', country=c('RU','CN','US','BRA','ESP','DEU','IND','ZAF'), start=1998, end=2011)
country <- datTOT1$country
datyear <- datTOT1$year
datmilnum <- datTOT1$MS.MIL.TOTL.P1
datpop <- datTOT2$SP.POP.TOTL

total_data = data.frame(country, datyear, datmilnum, datpop)
total_data <- total_data %>% mutate(milnumcapita = (datmilnum/datpop)*1000)
total_data$datyear <- as.numeric(total_data$datyear)
```

## Index Column
```{r}
nrow(total_data)
total_data$Index <- 0
for (i in 381:493) {
  total_data$Index[i - 381] <- i
  i + 1
}
```

## Write to Excel File
```{r}
openxlsx::addWorksheet(projectdata, "ArmedForces")
openxlsx::writeData(projectdata, "ArmedForces", total_data)
openxlsx::saveWorkbook(projectdata, "projectdata.xlsx", overwrite = TRUE)
```


## Plots

### Total Change for relevant countries over wide timeframe
```{r, cache=TRUE}
library(WDI)
library(tidyverse)
library(ggplot2)

dat = WDI(indicator='MS.MIL.TOTL.P1', country=c('RU','CN','US'), start=1998, end=2011)
dat2 = WDI(indicator='SP.POP.TOTL', country=c('RU','CN','US'), start=1998, end=2011)

country <- dat$country
datyear <- dat$year
datmilnum <- dat$MS.MIL.TOTL.P1
datpop <- dat2$SP.POP.TOTL
cleaned_data= data.frame(country, datyear, datmilnum, datpop)

cleaned_data <- cleaned_data %>% mutate(milnumcapita = (datmilnum/datpop)*1000)

ggplot(cleaned_data, aes(datyear, milnumcapita, color=country)) + geom_line() + 
  xlab('Year') + ylab('Number Active Mil. Personnel(per thousand)')
```

### Expected change in AF leading up to 2001
```{r, cache=TRUE}
dat = WDI(indicator='MS.MIL.TOTL.P1', country=c('RU','CN','US'), start=1998, end=2001)
dat2 = WDI(indicator='SP.POP.TOTL', country=c('RU','CN','US'), start=1998, end=2001)

country <- dat$country
datyear <- dat$year
datmilnum <- dat$MS.MIL.TOTL.P1
datpop <- dat2$SP.POP.TOTL
cleaned_data= data.frame(country, datyear, datmilnum, datpop)

cleaned_data <- cleaned_data %>% mutate(milnumcapita = (datmilnum/datpop)*1000) 
write.csv(cleaned_data, "ChangePre2001.csv", row.names=FALSE)

ggplot(cleaned_data, aes(datyear, milnumcapita, color=country)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE)+xlab('Year') + 
  ylab('Number Active Mil. Personnel(per thousand)')

cleaned_america = filter(cleaned_data, country=="United States")
cleaned_america
lm(milnumcapita~datyear,data=cleaned_america)
```

### Actual Change post-2001, both for 2004 and 2008(end of first and second Bush terms)
```{r, cache=TRUE}
dat = WDI(indicator='MS.MIL.TOTL.P1', country=c('RU','CN','US'), start=2001, end=2004)
dat2 = WDI(indicator='SP.POP.TOTL', country=c('RU','CN','US'), start=2001, end=2004)

country <- dat$country
datyear <- dat$year
datmilnum <- dat$MS.MIL.TOTL.P1
datpop <- dat2$SP.POP.TOTL
cleaned_data= data.frame(country, datyear, datmilnum, datpop)

cleaned_data <- cleaned_data %>% mutate(milnumcapita = (datmilnum/datpop)*1000)
write.csv(cleaned_data, "Change2004.csv", row.names=FALSE)

ggplot(cleaned_data, aes(datyear, milnumcapita, color=country)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE)+xlab('Year') + 
  ylab('Number Active Mil. Personnel(per thousand)')

cleaned_america = filter(cleaned_data, country=="United States")
cleaned_america
lm(milnumcapita~datyear,data=cleaned_america)

dat3 = WDI(indicator='MS.MIL.TOTL.P1', country=c('RU','CN','US'), start=2001, end=2008)
dat4 = WDI(indicator='SP.POP.TOTL', country=c('RU','CN','US'), start=2001, end=2008)

country <- dat3$country
datyear <- dat3$year
datmilnum <- dat3$MS.MIL.TOTL.P1
datpop <- dat4$SP.POP.TOTL
cleaned_data= data.frame(country, datyear, datmilnum, datpop)

cleaned_data <- cleaned_data %>% mutate(milnumcapita = (datmilnum/datpop)*1000)
write.csv(cleaned_data, "Change2008.csv", row.names=FALSE)

ggplot(cleaned_data, aes(datyear, milnumcapita, color=country)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE)+xlab('Year') + 
  ylab('Number Active Mil. Personnel(per thousand)')

cleaned_america = filter(cleaned_data, country=="United States")
cleaned_america
lm(milnumcapita~datyear,data=cleaned_america)
```

### Change in non-conflict nations
```{r, cache=TRUE}
dat = WDI::WDI(indicator='MS.MIL.TOTL.P1', country=c('BRA','ESP','DEU','IND','ZAF'), start=1998, end=2011)
dat2 = WDI::WDI(indicator='SP.POP.TOTL', country=c('BRA','ESP','DEU','IND','ZAF'), start=1998, end=2011)

country <- dat$country
datyear <- dat$year
datmilnum <- dat$MS.MIL.TOTL.P1
datpop <- dat2$SP.POP.TOTL
cleaned_data= data.frame(country, datyear, datmilnum, datpop)

cleaned_data <- cleaned_data %>% mutate(milnumcapita = (datmilnum/datpop)*1000)
write.csv(cleaned_data, "ChangeNonConflict.csv", row.names=FALSE)

ggplot(cleaned_data, aes(datyear, milnumcapita, color=country)) + geom_line() + 
  xlab('Year') + ylab('Number Active Mil. Personnel(per thousand)')
```

```{r}
DataExplorer::create_report(votes)
DataExplorer::create_report(gdp)
DataExplorer::create_report(total_data)
```

